\simname 2.0 User's Guide

### 2013-11-26 doc/manual.tex
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/backoff_alloc@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/pbooksim@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/systemsim_interface@5188 ce7aba0d-9f81-40a6-a495-cecae84b0515
update copyright notice
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/priority_tree@5188 ce7aba0d-9f81-40a6-a

% Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
% All rights reserved.
% 
% Redistribution and use in source and binary forms, with or without 
% modification, are permitted provided that the following conditions are met:
% 
% Redistributions of source code must retain the above copyright notice, this 
% list of conditions and the following disclaimer.
% Redistributions in binary form must reproduce the above copyright notice, 
% this list of conditions and the following disclaimer in the documentation 
% and/or other materials provided with the distribution.
% Neither the name of the Stanford University nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
% 
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" 
% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE 
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE 
% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE 
% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR 
% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF 
% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS 
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN 
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) 
% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE 
% POSSIBILITY OF SUCH DAMAGE.

@@

A note for Windows users:
The above instructions have been tested to work with the newest version of Cygwin.

@@

\begin{figure}
\begin{verbatim}
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
0%=================================
% Average latency = 10.3206
% Accepted packets = 0.113 at node 38 (avg = 0.149031)
lat(1) = 10.3206;
bins = [ 0 ...999];
freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
 49 10 7 1 0 ....

@@

% Warmed up ...Time used is 3000 cycles

@@

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6030 cycles
====== Traffic class 0 ======
Overall average latency = 10.2897 (1 samples)
Overall average accepted rate = 0.14834 (1 samples)
Overall min accepted rate = 0.129043 (1 samples)
Average hops = 4.99839 (57153 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

After the warm up periods have passed (default to 3$\times$sample\_period), the simulator prints the
``\texttt{Warmed up}'' message and resets all the simulation statistics.
Then, the measurement phase begins and statistics continue to be
reported after each sample period.  Once the measurement periods have
passed, all the measurement packets are drained from the network
before final latency and throughput numbers are reported.  Details of
the configuration parameters used to control the length of the
simulation phases are covered in Section~\ref{sec:sim_params}.

@@

// Routing

routing_function = dim_order;

// Flow control

num_vcs     = 8;
vc_buf_size = 8;

wait_for_tail_credit = 1;

// Router architecture

vc_allocator = islip;
sw_allocator = islip;
alloc_iters  = 1;

credit_delay   = 2;
routing_delay  = 1;
vc_alloc_delay = 1;

@@

// Traffic

traffic                = transpose;
const_flits_per_packet = 20;

// Simulation

sim_type       = latency;
injection_rate = 0.01;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file.

@@

\item[k] Network radix, the number of routers per dimension
\item[n] Network dimension
\item[c] Network concentration, the number of nodes sharing a single router. Typically set to 1, >1 only in networks that has concentration (i.e. cmesh)
\item[x] (NoC simulations only) The number of routers in the X dimension. Used to calculate channel latency between routers. 
\item[y] (NoC simulations only) The number of routers in the y dimension. Used to calculate channel latency between routers. 
\item[xr] (NoC simulations only) For networks that have c>1, the number of nodes in the x direction per router. Used to calculate channel latency between routers. 
\item[yr] (NoC simulations only) For networks that have c>1, the number of nodes in the y direction per router. Used to calculate channel latency between routers. 
\item[limit] Emulate partially filled networks. Nodes greater or equal to limit does not generate or receive packets during simulation. Though depending on the routing algorithm the routers associated with those nodes may be use to route passing by traffic.

@@

\begin{opt_list}{topologies}
\item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter
determines the network's dimension.

@@

\item[single] A network with a single node, used for testing single
router performance.  The number of input and output ports for the node
is determined by the \texttt{in\_ports} and \texttt{out\_ports} parameters,
respectively.

@@

@@

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.

@@

Both the \texttt{mesh} and \texttt{torus} topologies support the
addition of random link failures with the \texttt{link\_failures}
parameter.  The value of \texttt{link\_failures} determines the number
of channels that are randomly removed from the topology and are thus
no longer available for forwarding packets.  Moreover, the
randomization for failed channels is controlled by selecting an
integer value for the \texttt{fail\_seed} parameter --- a fixed seed
gives a fixed set of failed channels, independent of other
randomization in the simulation.  Also, note that only certain routing
functions support this feature (see Section~\ref{sec:routing_algs}).

@@

The \texttt{routing\_function} parameter selects a routing algorithm
for the topology.  Many routing algorithms need multiple virtual
channels for deadlock freedom (VCDF). In addition to \texttt{routefunc.cpp}, some topologies source files include additional routing functions. Also, the simulator code is structured so that additional routing algorithms can be added with minimal changes to the overall simulator (see the \texttt{routefunc.cpp} file in the simulator's source code).

@@

\item[output\_delay] The processing delay incurred in the output queue
of a router.
\end{opt_list}

@@

\item[filter\_spec\_grants] Determines how speculative grants are masked (\texttt{any\_nonspec\_gnts}: any non-speculative grant inhibits all speculative grants; \texttt{confl\_nonspec\_reqs}: speculative grants are inhibited by conflicting non-speculative requests; \texttt{confl\_nonspec\_gnts}: speculative grants are inhibited by conflicting non-speculative grants).

@@

@@

\item[sw\_alloc\_arb\_type] If the switch allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.

\item[sw\_alloc\_iters] If the switch allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed. This overrides the global \texttt{alloc\_iters} parameter.

@@

\item[vc\_alloc\_arb\_type] If the VC allocator is a separable input- or output-first allocator, this parameter selects the type of arbiter to use.

\item[vc\_alloc\_iters] If the VC allocator is of type \texttt{islip}, \texttt{pim} or \texttt{select}, this parameter controls how many allocation iterations should be performed. This overrides the global \texttt{alloc\_iters} parameter.

@@

\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network. This print out should be fed into noc\_viewer for a graphic display of the activity inside the network. Currently not working.

>>>>

% Copyright (c) 2007-2012, Trustees of The Leland Stanford Junior University
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without
% modification, are permitted provided that the following conditions are met:
%
% Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
% Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation
% and/or other materials provided with the distribution.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
% POSSIBILITY OF SUCH DAMAGE.

@@

A note for Windows users:
The above instructions have been tested to work with Cygwin 1.7.18.

@@

\begin{figure}
\begin{verbatim}
BEGIN Configuration File: examples/torus88

@@

// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 2;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File: examples/torus88
Class 0:
Packet latency average = 33.2807
	minimum = 7
	maximum = 73

@@

Warmed up ...Time used is 2000 cycles

Draining all recorded packets ...
Draining remaining packets ...
Time taken is 5104 cycles
====== Overall Traffic Statistics ======
====== Traffic class 0 ======
Packet latency average = 33.6964 (1 samples)
	minimum = 7 (1 samples)
	maximum = 84 (1 samples)

Total run time 1.01837
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

After the warm up periods have passed (default to 3$\times$sample\_period), the simulator prints the
``\texttt{Warmed up}'' message and resets all the simulation statistics.
Then, the measurement phase begins and statistics continue to be
reported after each sample period.  The measurement phase 
typically last for last least 3 sample periods. 
Once the measurement periods have
passed, all the measurement packets are drained from the network
before final latency and throughput numbers are reported.  Details of
the configuration parameters used to control the length of the
simulation phases are covered in Section~\ref{sec:sim_params}.

@@

// Routing
routing_function = dor;

// Flow control
num_vcs     = 8;
vc_buf_size = 8;
wait_for_tail_credit = 1;

// Router architecture
vc_allocator = islip;
sw_allocator = islip;
alloc_iters  = 1;

credit_delay   = 2;
routing_delay  = 0;
vc_alloc_delay = 1;
sw_alloc_delay = 1;

@@

// Traffic
traffic = transpose;
packet_size = 20;

// Simulation
sim_type = latency;

injection_rate = 0.005;

\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in 
Section~\ref{sec:run_example}.  This section lists the major
configuration parameters for the simulator. Additional description can be found in the 
configuration parameter class file \texttt{booksim\_config.cpp}.
A user can incorporate additional options by changing the this file.

@@

\item[k] Network radix, the number of routers per dimension
\item[n] Network dimension
\item[c] Network concentration, the number of nodes sharing a single
  router. Typically set to 1, >1 only in networks that has
  concentration (i.e. cmesh).
\item[x] (NoC simulations only) The number of routers in the X
  dimension. Used to calculate channel latency between routers.  
\item[y] (NoC simulations only) The number of routers in the y
  dimension. Used to calculate channel latency between routers.  
\item[xr] (NoC simulations only) For networks that have c>1, the
  number of nodes in the x direction per router. Used to calculate
  channel latency between routers.  
\item[yr] (NoC simulations only) For networks that have c>1, the
  number of nodes in the y direction per router. Used to calculate
  channel latency between routers.

@@

\begin{opt_list}{topologies}
\item[fly] A $k$-ary $n$-fly (butterfly) topology. The \texttt{k}
parameter determines the network's radix and the \texttt{n} parameter
determines the network's dimension. Note: a $k$-ary 1-fly is
essentially a single radix-k router, useful for testing.

@@

@@

\item[dragonfly] A topology based on the paper ``Technology-driven,
  highly-scalable dragonfly topology.'' ISCA 2008

@@

\item[anynet] A topology based on an user input file specifying
  connectivity of nodes and routers.

@@

%Both the \texttt{mesh} and \texttt{torus} topologies support the
%addition of random link failures with the \texttt{link\_failures}
%parameter.  The value of \texttt{link\_failures} determines the number
%of channels that are randomly removed from the topology and are thus
%no longer available for forwarding packets.  Moreover, the
%randomization for failed channels is controlled by selecting an
%integer value for the \texttt{fail\_seed} parameter --- a fixed seed
%gives a fixed set of failed channels, independent of other
%randomization in the simulation.  Also, note that only certain routing
%functions support this feature (see Section~\ref{sec:routing_algs}).

@@

The \texttt{routing\_function} parameter selects a routing algorithm
for the topology.  Many routing algorithms need multiple virtual
channels for deadlock freedom. In addition to \texttt{routefunc.cpp},
some topologies source files include additional routing
functions. Also, the simulator code is structured so that additional
routing algorithms can be added with minimal changes to the overall
simulator (see the \texttt{routefunc.cpp} file in the simulator's
source code).

@@

%\item[output\_delay] The processing delay incurred in the output queue
%of a router.
\end{opt_list}

@@

%\item[filter\_spec\_grants] Determines how speculative grants are masked (\texttt{any\_nonspec\_gnts}: any non-speculative grant inhibits all speculative grants; \texttt{confl\_nonspec\_reqs}: speculative grants are inhibited by conflicting non-speculative requests; \texttt{confl\_nonspec\_gnts}: speculative grants are inhibited by conflicting non-speculative grants).

@@

\item[arb\_type] If the VC or switch  allocator is a separable
  input- or output-first allocator, this parameter selects the type of
  arbiter to use.

@@

@@

@@

%\item[viewer\_trace] The simulator will generate very verbose print out of all activity inside the network. This print out should be fed into noc\_viewer for a graphic display of the activity inside the network. Currently not working.
